### 2024-01-06 texts/paper_closing.tex
re-organizing for EGOS

% Created 2024-01-06 za 15:19
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}

@@

\section*{Private museum closures}

\subsection*{Abstract: TODO}

@@

In recent years, private, collector-run museums have seen an unprecedented proliferation.
These museums, run by individual collectors, constitute an addition to established organizational forms of art provision such as museums run/financed by the state or foundations and associations.
Despite their occasionally spectacular/sensationalist press coverage (TODO:ref), these institutions and the motivations of their founders have remained remarkably intransparent.
Financed in large parts by their founders' private wealth, few reporting requirements exists; and while some founders highlight their charitable and distintersted intentions in media coverage, such statements intended for a public audience might not necessarily accurately reflect underlying founder motivations.

@@

@@

\subsection*{Theory}

TODO: the theory section (including the hypotheses) is mostly copied from an earlier literature review on organizational closure, and while there are some references to how it could relate to private museums, it isn't integrated with the current focus on collector decision making and private museum sustainability.

\subsubsection*{Size}

Larger NPOs are generally found to have lower mortality rates \parencite{Hager_1999_demise,Duckles_Hager_Galaskiewicz_2005_close}, with size being measured through indicators such as assets \parencite{Lu_Shon_Zhang_2019_dissolution}, number of employees \parencite{Bielefeld_1994_survival} and volunteers \parencite{Fernandez_2007_dissolution}, or revenues \parencite{Searing_2020_zombies}.
Smaller organiations are argued to face greater difficulty in collecting resources and recruiting staff as well as relatively higher costs for administration and regulation \parencite{Helmig_Ingerfurth_Pinz_2013_nonprofit}, while being less able to diversify or benefit from economies of scale \parencite{Hager_1999_demise}. 
Larger organizations are also argued to be able to waste some resources during normal times which can be utilized effectively during situations of crisis without having to scale back services (which generate revenue) \parencite{Hager_2001_vulnerability} and hence avoid downsizing, which can lead to dissolution \parencite{Duckles_Hager_Galaskiewicz_2005_close,Bielefeld_1994_survival}.

\bigbreak
\noindent
\textbf{H\textsubscript{size}}: Larger private museums are less likely to close than smaller ones.

\subsubsection*{Resource diversification}

A substantial amount of literature focuses on resources as resource depletion is a leading cause of (nonprofit) organizational demise. 
Diversification (i.e. a multitude of funding sources) is found to reduce mortality \parencite{Fernandez_2007_dissolution,Bielefeld_1994_survival,Hager_2001_vulnerability,Lu_Shon_Zhang_2019_dissolution}; similarly volatility (rapid changes in revenues) is found to increase mortality \parencite{Mayer_2022_slimmer}. 
While private museums are likely already generally leaning toward higher concentration than NPOs generally due to high reliance on the financial contributions of the founder-collector, there might still be variation in the degree of revenue concentration as museums might vary in the proportion of revenues received from ticket prices, restaurants, cafes, museum shops or other activities.

\bigbreak
\noindent
\textbf{H\textsubscript{diversification}}: Private museums with more diversified revenue channels are less likely to close than more diversified ones.

@@

\subsubsection*{legitimacy/understandability}

Legitimacy is necessary for being understood/tolerated/supported by other the organizations in the environemnt; to the extent that this influences resource acquisition it could influence survival prospects.
While \textcite{Bielefeld_1994_survival} finds that NPOs who pursue less legitimation strategies (obtaining endorsements, lobbying or contributing to local causes) are more likely to close, other studies find limited impact of legitimacy on survival, both when measured via density \parencite{Bogaert_etal_2014_ecological} or from archival sources and interviews \parencite{Fernandez_2007_dissolution}.
In the context of private museums, decline of legitimacy/status could lead to lower visitor numbers, less discounts from art dealers, less government grants (?) or lower chances on collaborations with other museums or corporations, resulting in lower revenues and higher acquisition costs.

Legitimacy might be obtained by isomorphism, i.e. adopting features associated with blueprints \parencite{diMaggio_1983_iron}.
Understandability/sense-making \parencite{Glynn_Abzug_2002_names} might thus be limited if an organization is atypical \parencite{Rosch_1975_family}, i.e. displays an unusual combination of features.
In the case of private museums, such cases might be combinations of features from house museums (e.g. accessibility only by appointment) with full-service museums (e.g. extensive activities). 
Next to "hard" organizational features, adherence to naming conventions has been found to enhance legitimacy \parencite{Glynn_Abzug_2002_names}; organizations that adhere to field norms about name length, name ambiguity (usage of artificial names) and name domain specificity (mentioning industry) were judged more legitimate. 
However, the category of museums could be relatively flexible (for example, it is generally not subject to state regulation), which might result in a relatively high tolerance of diversity as "anything goes" and hence limited "devaluation" of non-conforming/atypical members.

\bigbreak
\noindent
\textbf{H\textsubscript{typicality}}: Atypical private museums have higher mortality rates.

Legitimacy might also be measured by inclusion in industry registers, such as the Museum of the World (MOW) database \parencite{deGruyter_2021_MOW}.

\bigbreak
\noindent
\textbf{H\textsubscript{mow}}: Private museums included in the MOW database have higher survival chances.

\subsubsection*{reputation}

Reputation has been argued to enhance organizational survival \parencite{Rao_1994_reputation}.
However, reputation effects could also be a spurious effect of organizational capabilities as more competent organizations are more able to both aquire reputation through certification contests and survive in long run \parencite{Goldfarb_Zavyalova_Pillai_2018_Rao}.
Recognition might be measured by inclusion in the Artnews top 200 collector ranking.

\bigbreak
\noindent
\textbf{H\textsubscript{artnews}}: Private museums of collectors included in the Artnews top 200 collector ranking are less likely to close.

@@

\subsubsection*{Predictors}

TODO: the current do not yet measure all the concepts discussed in the theory, and some are not justified yet on theoretical grouns (founder name in museum name, gender, density).

\bigbreak
\noindent
On the level of the founder, gender constitutes a time-invariant variable; with possible values being male, female and couple.
Founder death is constructed as a time-varying covariate, and is set to 1 for museum-years after the death of the founder.

On the museum level, self-identification measures whether a the name of a museum includes the term "Museum", "Foundation", "Collection", or other (such as "Villa", "Institute", "Center" or "Gallery", which are grouped together as they are less frequent).
Further analyzing the name, a binary indicator also measures whether the name of the founder is included.
Finally, a binary predictor measures whether the museum is included in the "Museums of the World" database \parencite{deGruyter_2021_MOW}, a database containig 55k museums.
On the level of the environment, the density (number) of private museum per country is included both in linear and squared form.
Organizational theory \parencite{hannan89_organ} predicts a U-shaped relationship between density and mortality: While at low levels, increases in the number of organizations are argued to primarily increase legitimacy/familiarity and lower exit rates, increases at higher numbers increase the competition over resources and thereby raise mortality rates.

\bigbreak
\noindent
TODO: 
\begin{itemize}
\item add museum size/diversification indicators: most museum-level variables that we collected have many missing values (figure \ref{fig:p_vrblcvrg_ratio}), so it's challenging to get good estimates out of them. I'm thinking of using some dimensionality reduction like PCA.
\item add inclusion in Artnews collector ranking
\item add Tax incentives (country-level)
\item expand densities: 
\begin{itemize}
\item per capita
\item include MOW museums, for now density is based only on private museums
\item within proximate environment (circle of X kilometer around museum), rather than country-level
\end{itemize}
\item account for clustering within countries/regions
\end{itemize}

@@

In a second step, I use Cox proportional hazard models which estimate the relationship between covariates and the hazard rate.
The Cox-proportional hazard model has the form \(h(t,\mathbf{x}) = h_0(t) \psi\), with \(\psi = \exp(\sum_{j=1}^p x_j \beta_j)\), with \(h(t)\) as the hazard rate, \(h_0(t)\) as the baseline hazard rate, \(\mathbf{x}\) as the set of variables indexed by \(j\), and \(\beta\) as the coefficient of the \(j\) -th variable.
Estimation via partial maximum likelihood allows estimation of \(\beta\) without estimating the baseline hazard rate, which is therefore not constrained to a parametric form. 
Coefficients are interpeted as log hazard ratios, i.e. as multipliers of the baseline hazard rate.

This model assume proportional hazards, i.e. that the coefficient of a variable stays constant over time, which indicates that for a variable at every time point the hazard ratio is the same.
So far, tests show a violation of the proportionality assumption only for the variable indicating whether the museum name includes the name of the founder  which furthermore is marginal as it slightly varies by the type of time transformation (table \ref{tbl:t_coxzph}, figure \ref{fig:p_coxphdiag_more}).

@@

Figure \ref{fig:p_hazard} shows the hazard rate over time, figure \ref{fig:p_surv} the corresponding survival probability.
The hazard rate doubles over the first 10 years from approximately 0.45\% to 0.9\%, at which value it approximately stays constant for the next 20 years, after which it appears to decline.
However, the decline after thirty years is presumably less certain as very few private museums have already reached this age (figure \ref{fig:p_agedens}).

@@

\begin{sloppypar}
\printbibliography
\end{sloppypar}

\section*{Appendix}

\begin{figure}[htbp]
\centering
\includegraphics[width=19cm]{../figures/p_yeardens.pdf}
\caption{\label{fig:p_yeardens}Private Museum-year Distribution}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=17cm]{../figures/p_agedens.pdf}
\caption{\label{fig:p_agedens}Age distribution of private museums}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=19cm]{../figures/p_coxphdiag_more.pdf}
\caption{\label{fig:p_coxphdiag_more}Model Diagnostics}
\end{figure}

% latex table generated in R 4.3.2 by xtable 1.8-4 package
% Fri Dec  8 22:31:01 2023
\begin{table}[ht]
\centering
\begin{tabular}{p{0mm}lD{.}{.}{5}D{.}{.}{5}D{.}{.}{5}}
  \hline 
 \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{Kaplan-Meier} & \multicolumn{1}{l}{Identity} & \multicolumn{1}{l}{Rank}\\ 
 \hline
  \multicolumn{2}{l}{\textbf{Founder}} \\ 
 & Founder died & .59 & .55 & .72 \\ 
   & Gender - Couple & .31 & .23 & .32 \\ 
   & Gender - Female & .61 & .68 & .56 \\ 
   \multicolumn{2}{l}{\textbf{Museum}} \\ 
 & AN Ranking - Dropped & .88 & .98 & .71 \\ 
   & AN Ranking - Included & .59 & .46 & .41 \\ 
   & Founder name in Museum name & .09 & .09 & .08 \\ 
   & MOW inclusion & .28 & .28 & .51 \\ 
   & Self-Identification - Collection & .60 & .53 & .36 \\ 
   & Self-Identification - Foundation & .93 & .93 & .84 \\ 
   & Self-Identification - Other & .90 & .83 & .85 \\ 
   \multicolumn{2}{l}{\textbf{Environment}} \\ 
 & PM density & .95 & 1.00 & .90 \\ 
   & PM density$^{2}$ & .77 & .70 & .78 \\ 
   \multicolumn{2}{l}{\textbf{Miscellaneous}} \\ 
 & Global & .52 & .39 & .51 \\ 
   \hline 
 \multicolumn{3}{l}{\footnotesize{\textsuperscript{***}p $<$ 0.001;\textsuperscript{**}p $<$ 0.01;\textsuperscript{*}p $<$ 0.05.}}
\end{tabular}
\caption{Z-test of proportional hazards} 
\label{tbl:t_coxzph}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics[width=18cm]{../figures/p_vrblcvrg_ratio.pdf}
\caption{\label{fig:p_vrblcvrg_ratio}PMDB variable coverage (abs/rel prop) by museum status and variable group}
\end{figure}
\end{document}

>>>>

% Created 2024-01-06 za 18:00
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}

@@

\section*{The closure of private museums: Legitimacy, identity and reputation}

@@

In recent years, private collector-run museums have seen an unprecedented proliferation.
These museums, run by individual collectors, constitute an addition to established organizational forms of art provision such as museums operated and financed by the state or foundations and associations.
While the emergence of this organizational form has attracted attention (cf. \cite{Kolbe_etal_2022_privatemuseum} for a literature review), the process of closure has received considerably less attention.
In this paper, I investigate

argue that studying closures does not only provide insights into private museums, but into more general processes of organizational identity, legitimation and reputation.

In particular concerning identity, I find that museums which are recognized by an international museum database are less likely to close than those which are not.
With regards to legitimation, results indicate that that museums with more ambiguous names are more likely to close.
Finally, a reputation effect is indicated by the fact that museums whose founders are dropped from a collector ranking are more likely to close than museums whose founders are included.

\subsection*{Contributions}

\begin{itemize}
\item Intransparent founder motivations
\end{itemize}

Despite their occasionally spectacular/sensationalist press coverage, these institutions and the motivations of their founders have remained remarkably intransparent.
Financed in large parts by their founders' private wealth, few reporting requirements exists; and while some founders highlight their charitable and distintersted intentions in media coverage, such statements intended for a public audience might not necessarily accurately reflect underlying founder motivations.

@@

\begin{itemize}
\item comparison to other forms (more descriptively)
\end{itemize}

@@

\subsection*{Theoretical framework}

\subsubsection*{Legitimacy and the survival prospects of non-profit organizations}

Legitimacy is necessary for being understood, tolerated and/or supported by other the organizations in the environemnt; to the extent that this influences resource acquisition it could influence survival prospects.
While \textcite{Bielefeld_1994_survival} finds that NPOs who pursue less legitimation strategies (obtaining endorsements, lobbying or contributing to local causes) are more likely to close, other studies find limited impact of legitimacy on survival, both when measured via density \parencite{Bogaert_etal_2014_ecological} or from archival sources and interviews \parencite{Fernandez_2007_dissolution}.
In the context of private museums, decline of legitimacy/status could lead to lower visitor numbers, less discounts from art dealers, less government grants or lower chances on collaborations with other museums or corporations, resulting in lower revenues and higher acquisition costs.

Legitimacy might be obtained by isomorphism, i.e. adopting features associated with blueprints \parencite{diMaggio_1983_iron}.
Understandability \parencite{Glynn_Abzug_2002_names} might thus be limited if an organization is atypical \parencite{Rosch_1975_family}, i.e. displays an unusual combination of features.
In the case of private museums, such cases might be combinations of features from house museums (e.g. accessibility only by appointment) with full-service museums (e.g. extensive activities). 
Next to "hard" organizational features, adherence to naming conventions has been found to enhance legitimacy \parencite{Glynn_Abzug_2002_names}; organizations that adhere to field norms about name length, name ambiguity (usage of artificial names) and name domain specificity (mentioning industry) were judged more legitimate. 
However, the category of museums could be relatively flexible (for example, it is generally not subject to state regulation), which might result in a relatively high tolerance of diversity as "anything goes" and hence limited "devaluation" of non-conforming, atypical members \parencite{Bogaert_etal_2014_ecological}.

\bigbreak
\noindent
\textbf{Hypothesis 1}: Private museums which are considered more legitimate are less likely to close.

\subsubsection*{Reputation}

Reputation has been argued to enhance organizational survival \parencite{Rao_1994_reputation}.
However, reputation effects could also be a spurious effect of organizational capabilities as more competent organizations are more able to both aquire reputation through certification contests and survive in long run \parencite{Goldfarb_Zavyalova_Pillai_2018_Rao}.
\bigbreak
\noindent
\textbf{H\textsubscript{artnews}}: Private museums of collectors included in the Artnews top 200 collector ranking are less likely to close.

@@

@@

\subsubsection*{Legitimacy}

Two measures are constructed that aim to capture an effect of legitimacy on closing.
First, self-identification measures whether a the name of a museum includes the term "Museum", "Foundation", "Collection", or other (such as "Villa", "Institute", "Center" or "Gallery", which are grouped together as they are less frequent).
Secondly, binary predictor measures whether the museum is included in the "Museums of the World" database \parencite{deGruyter_2021_MOW}, a database containig 55k museums.

\subsubsection*{Reputation}

Recognition is measured via inclusion of the founder in the Artnews top 200 collector ranking \parencite{Artnews_ranking}.
This ranking, established in 1990, lists the 200 collectors the editors of the magazine consider most relevant.
For each year, a museums can have one of three values: "Not included"; in which case the founder is not included in the year in question and has not been so in the past, "Included"; when the founder is included, or "Dropped", for cases where the founder was included at some point in the past but is no longer included in the current year.
Cases where multiple individuals are involved in the founding of a private museum are resolved by aggregating the Artnews ranking inclusion history to the founder founder level (i.e., if one of two founders is dropped from the Artnews ranking but the other remains included, the founder couple as a whole is considered included).

\subsubsection*{Control variables}

\bigbreak
\noindent
A number of control variables are used: 
On the level of the founder, gender constitutes a time-invariant variable; with possible values being male, female and couple.
Founder death is constructed as a time-varying covariate, and is set to 1 for museum-years after the death of the founder. 
Further analyzing the name, a binary indicator also measures whether the name of the founder is included in the name of the museum as a measure of centrality of the founder's vision.
Organizational theory \parencite{hannan89_organ} predicts a U-shaped relationship between density and mortality:
While at low levels, increases in the number of organizations are argued to primarily increase legitimacy and lower exit rates, increases at higher numbers increase the competition over resources and thereby raise mortality rates.
On the level of the environment, therefore the density (number) of private museum per country is included both in linear and squared form.

@@

In a second step, we use Cox proportional hazard models which estimate the relationship between covariates and the hazard rate.
The Cox-proportional hazard model has the form \(h(t,\mathbf{x}) = h_0(t) \psi\), with \(\psi = \exp(\sum_{j=1}^p x_j \beta_j)\), with \(h(t)\) as the hazard rate, \(h_0(t)\) as the baseline hazard rate, \(\mathbf{x}\) as the set of variables indexed by \(j\), and \(\beta\) as the coefficient of the \(j\) -th variable.
Estimation via partial maximum likelihood allows estimation of \(\beta\) without estimating the baseline hazard rate, which is therefore not constrained to a parametric form. 
Coefficients are interpeted as log hazard ratios, i.e. as multipliers of the baseline hazard rate.

@@

Figure \ref{fig:p_hazard} shows the hazard rate over time, figure \ref{fig:p_surv} the corresponding survival probability.
The hazard rate doubles over the first 10 years from approximately 0.45\% to 0.9\%, at which value it approximately stays constant for the next 20 years, after which it appears to decline.
However, the decline after thirty years is presumably less certain as very few private museums have already reached this age.

@@

\begin{sloppypar}
\printbibliography
\end{sloppypar}
\end{document}
